\batcave: Calling somatic mutations with a tumor- and site-specific prior

### 2019-12-09 NAR_DRAFT.tex
add language to remove folding

\subsection{Estimation of the mutation profile}
Many aspects of genomic architecture can affect the somatic mutation rate at multiple scales \cite{Buisson2019}.
Here we focus on a small-scale feature, the trinucleotide context, which is known to strongly affect the prior probability of single-nucleotide mutation \cite{Nik-Zainal2012a,Alexandrov2015,Lee-Six2018}.
The trinucleotide context of a genomic site consists of the identity of the reference base and the 3' and 5' flanking bases.
Folding the central base to the pyrimidines, there are two possible bases at the focal site, and there are four possible bases 3' and 5' of the focal site, yielding $2 \cdot 4 \cdot 4$ possible tri-nucleotide contexts $C$.
At the focal site, a mutation $m$ can be to any of three alternate alleles.
Indexing by the $c=\{1 \dots 32\}$ contexts and by the $m = \{1 \dots 3\}$ alternate bases, we have 96 possible substitution types $S_{m,c}$.
Eq.~\ref{eqn:full_prior} is then
\begin{equation}
  \label{eqn:detailed_prior}
  \mathrm{p}(\mathrm{S}_{m,c}) = \mathrm{p}(m \mid C = c) \mathrm{p}(C = c \mid M)\frac{\mathrm{p}(M)}{\mathrm{p}(C = c)}.
\end{equation}
The first two terms on the right-hand side can be estimated from the observed mutation profile (Fig.~\ref{NAR-realsigfig}).
%Here $\mathrm{p}(\mathrm{S}_{m,c})$ represents probability that a substitution of type $\mathrm{S}_{m,c}$ will occur.

>>>>

\subsection{Estimation of the mutation profile}
Many aspects of genomic architecture can affect the somatic mutation rate at multiple scales \cite{Buisson2019}.
Here we focus on a small-scale feature, the trinucleotide context, which is known to strongly affect the prior probability of single-nucleotide mutation \cite{Nik-Zainal2012a,Alexandrov2015,Lee-Six2018}.
The trinucleotide context of a genomic site consists of the identity of the reference base and the 3' and 5' flanking bases.
\bkmcomment{There are six classes of base substitution -- C$>$A, C$>$G, C$>$T, T$>$A, T$>$C, T$>$G (all substitutions are referred to by the pyrimidine of the mutated Watson–Crick base pair).
This results in a total of 32 ($4\times 2 \times 4$) reference contexts and 3 alternate bases per reference base.}
%Folding the central base to the pyrimidines, there are two possible bases at the focal site, and there are four possible bases 3' and 5' of the focal site, yielding $2 \cdot 4 \cdot 4$ possible tri-nucleotide contexts $C$.
%At the focal site, a mutation $m$ can be to any of three alternate alleles.
Indexing by the $c=\{1 \dots 32\}$ contexts and by the $m = \{1 \dots 3\}$ alternate bases, we have 96 possible substitution types $S_{m,c}$.
Eq.~\ref{eqn:full_prior} is then
\begin{equation}
  \label{eqn:detailed_prior}
  \mathrm{p}(\mathrm{S}_{m,c}) = \mathrm{p}(m \mid C = c) \mathrm{p}(C = c \mid M)\frac{\mathrm{p}(M)}{\mathrm{p}(C = c)}.
\end{equation}
The first two terms on the right-hand side can be estimated from the observed mutation profile (Fig.~\ref{NAR-realsigfig}).
%Here $\mathrm{p}(\mathrm{S}_{m,c})$ represents probability that a substitution of type $\mathrm{S}_{m,c}$ will occur.
